---
title: "Data analysis"
author: "Aoyue Bi"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: simplex 
    highlight: haddock 
---

## work1: hapResult hapSummary hapNet

```{r}
library(geneHapR)
library(conflicted)
library(dplyr)
library(tidyverse)

conflict_prefer("select", "dplyr")
conflicts_prefer(dplyr::filter)
conflicted::conflicts_prefer(dplyr::rename)

target_genes <- c("PRDM16")
target_genes <- c("PRKCH")
target_genes <- c("CSGALNACT2")
target_genes <- c("CH25H")
###============================================================================= parameter 
popfile <- "/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/007_ASMai/000_popInfo/003_PopInfo_APG_addArchaic.txt"
## Read population information file (only needs to be read once)
dfpop <- read_tsv(popfile) %>% select(Assembly_ID, Pop, Anc)

dfpara <- readxl::read_xlsx("/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/011_adaptiveIntrogression/005_geneHaplotype/000_gene_hsmap/gene_block_hsmap.xlsx")

# 定义基础路径
# popfile <- "/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/007_ASMai/000_popInfo/003_PopInfo_APG_addArchaic.txt"
base_hapmap_path <- "/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/011_adaptiveIntrogression/001_DenOri"
base_output_path <- "/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/012_haplotype_network/001_haplotype"
###============================================================================= parameter end

df_target <- dfpara %>% filter(GeneName %in% target_genes)
dfpara0 <- df_target

for(i in 1:nrow(dfpara0)) {
  
  geneName <- dfpara0$GeneName[i]
  DenBlock <- str_sub(dfpara0$ChunkID[i],2) 
  archaic <- dfpara0$Archaic[i]
  
  Start_heatmap <- dfpara0$Start_heatmap
  End_heatmap <- dfpara0$End_heatmap
  
  cat("Processing gene:", geneName, ", chunk:", DenBlock, "\n")
  
  if(archaic == "Denisovan"){
    hapmapfile <- file.path(base_hapmap_path, paste0("Den_", DenBlock, ".merge.easy_region.filt_chr.hapmap.xlsx"))
    outfile_hapResult <- file.path(base_output_path, paste0(geneName,"_D",DenBlock, "_forhapResult.txt"))
    outfile_hapSummary <- file.path(base_output_path, paste0(geneName,"_D",DenBlock, "_forhapSummary.txt"))
    
  }
  if(archaic == "Neanderthal"){
    hapmapfile <- file.path(base_hapmap_path, paste0("Nean_", DenBlock, ".merge.easy_region.filt_chr.hapmap.xlsx"))
    outfile_hapResult <- file.path(base_output_path, paste0(geneName,"_N",DenBlock, "_forhapResult.txt"))
    outfile_hapSummary <- file.path(base_output_path, paste0(geneName,"_N",DenBlock, "_forhapSummary.txt"))
  }
  
  if(!file.exists(hapmapfile)) {
    cat(" File does not exist, skipping:", hapmapfile, "\n")
    next  # Skip current loop, continue to next gene
  }
  
  cat("File exists, starting processing...\n")  
  
  ### ============================== step1: Process raw file to obtain geno dataframe
  df0 <- read_tsv(hapmapfile) %>% 
    rename(Den = `Den-Den`, NeanAltai = `Nean-Altai`, NeanVin = `Nean-Vin`, NeanCha = `Nean-Cha`) %>% 
    arrange(POS)
  
  ### Data filtering, step 1: Filter POS based on conditions
  df1 <- df0 %>% 
    filter(str_length(REF) == 1, str_length(ALT) ==1) %>%
    select(-POS_HG38) %>% 
    pivot_longer(cols = `KOR01-Mat`:last_col(), names_to = "Assembly_ID", values_to = "Geno") %>% 
    left_join(dfpop,by=c("Assembly_ID")) %>% 
    ### Filter based on frequency, requiring the frequency of each position across all individuals to be greater than 0.05
    group_by(POS) %>%
    summarise(geno_sum = sum(Geno), n = 2*n(), fre = geno_sum/n) %>%
    filter(0.05 < fre, fre < 0.95 ) %>%
    ungroup() %>%
    select(POS)
  
  df2 <- df0 %>% 
    inner_join(df1,by = c("POS"))
  
  ### Data filtering, step 2: Filter samples
  df3 <- df2 %>% 
    pivot_longer(cols = `KOR01-Mat`:last_col(), names_to = "Assembly_ID", values_to = "Geno") %>% 
    left_join(dfpop,by=c("Assembly_ID")) %>% 
    # filter(Anc != "AFR") %>% 
    select(-c(Pop,Anc)) %>% 
    ### Transpose back to the original format
    pivot_wider(names_from = Assembly_ID, values_from = Geno)
  
  ### After filtering, prepare the file format required for pheatmap
  if(!is.na(Start_heatmap)){
    if(archaic == "Denisovan"){
      df4 <- df3 %>% 
        select(CHR = `#CHROM`, POS, REF, ALT, POS_HG38,c(Den:last_col())) %>% 
        ### POS is the coordinate in CHM13
        filter(POS >= Start_heatmap, POS <= End_heatmap) 
    }else {
      
      df4 <- df3 %>% 
        select(CHR = `#CHROM`, POS, REF, ALT, POS_HG38,c(NeanAltai,`KOR01-Mat`:last_col())) %>% 
        ### POS is the coordinate in CHM13
        filter(POS >= Start_heatmap, POS <= End_heatmap) 
    }
    
  }else{
    if(archaic == "Denisovan"){
      df4 <- df3 %>% 
        select(CHR = `#CHROM`, POS, REF, ALT, POS_HG38,c(Den:last_col()))
    }else{
      df4 <- df3 %>% 
        select(CHR = `#CHROM`, POS, REF, ALT, POS_HG38,c(NeanAltai,`KOR01-Mat`:last_col()))
    }
  }
  
  
  ### write_tsv(df4,"~/Documents/geon.txt" )
  ### step1: import data
  df <- df4 %>% mutate(across(6:last_col(), ~ case_when(
    .x == 0 ~ REF,
    .x == 1 ~ ALT,
    .x == 2 ~ ALT,
    TRUE ~ "N"  
  )))
  
  ################################################################################################ Filter POS
  ### ==============Reduce the number of SNPs to construct haplotype network with a smaller dataset
  # dfintro <- read_tsv("/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/011_adaptiveIntrogression/005_geneHaplotype/001_gene/002_introgressionSite.txt")
  # Get column names and create dataframe
  # df <- df %>% 
  #   filter(POS_HG38 %in% dfintro$POS_HG38)
  ### ==============end
  ### write_tsv(df,"~/Documents/geon_ATCG.txt" )
  ####============================================================================================ delimiter

  
  ### ============================== step2: Build gt.geno AccINFO
  gt.geno <- as.data.frame(df)
  
  #### Annotation information
  dfpop <- read_tsv("/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/007_ASMai/000_popInfo/003_PopInfo_APG_addArchaic.txt") %>% 
    select(Assembly_ID, Anc)
  
  AccINFO0 <- tibble(id = colnames(df)) %>% dplyr::slice(6:n()) %>% left_join(dfpop,by=c("id" = "Assembly_ID")) 
  AccINFO <- AccINFO0 %>% column_to_rownames("id")
  ### write_tsv(AccINFO0,"~/Documents/pop_group.txt" )
  
  # Import accession group/location information
  # AccINFO <- import_AccINFO("~/Documents/t.txt")
  ################################################################################ Construct objects hapResult and hapSummary
  ### step2: table2hap
  hapResult <- table2hap(gt.geno, hapPrefix = "H",
                         hetero_remove = TRUE, na_drop = TRUE) 
  # write_tsv(hapResult,"/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/012_haplotype_network/001_test/001_hapResult.txt")
  
  write_tsv(hapResult,outfile_hapResult)

  
  ### View haplotype results
  hapSummary <- geneHapR::hap_summary(hapResult)
  # write_tsv(hapSummary,"/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/012_haplotype_network/001_test/001_hapSummary.txt")
  write_tsv(hapSummary,outfile_hapSummary)
  
  ### Get frequency table for each haplotype
  ### df_hap_fre <- tibble(Hap = hapSummary$Hap, Fre = hapSummary$freq) %>% filter(str_starts(Hap,"H"))
  ### Get the haplotype for each sample
  df_sm_hap <- tibble(Hap = hapResult$Hap, Accession = hapResult$Accession) %>% filter(str_starts(Hap,"H"))
  write_tsv(df_sm_hap,str_replace(outfile_hapResult,".txt","_sm2hap_hsmap.txt"))
  
  ### Get frequency table for each haplotype
  df_sm_hap2 <- df_sm_hap %>% group_by(Hap) %>% dplyr::count()
  write_tsv(df_sm_hap2,str_replace(outfile_hapSummary,".txt","_hap2fre_hsmap.txt"))
  
  ##write_tsv(df_sm_hap,"/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/012_haplotype_network/001_haplotype/TRPV1_sm2hap_hsmap.txt")
  
  
  ################################################################################ build hapNet
  ### Generate haplonet
  hapNet <- get_hapNet(hapSummary,
                       AccINFO = AccINFO,
                       groupName = "Anc")
  
  ### cat hapNet
  #print.default(hapNet)
  
  hap_group_table <- attr(hapNet, "hapGroup")
  df_hap_group_table <- hap_group_table %>% as.data.frame() %>% pivot_wider(names_from = group,values_from = Freq)
  write_tsv(df_hap_group_table, str_replace(outfile_hapSummary,".txt","_hap2group.txt"))
  
  # Start redirecting output to file
  sink(str_replace(outfile_hapSummary,".txt","_hapNet_output.txt"))
  print.default(hapNet)
  sink()  # End redirection
  
  
} ### for

```

## work2: Calculate pairwise differences between haplotypes - without filtering haplotypes

Output results: difference table and heatmap

write_csv(diff_table, file.path(base_output_path,str_c(geneName,"\_",DenBlock,"\_haplotype_pairwise_differences.csv")))

```{r}
library(tidyverse)
library(reshape2)
library(ggplot2)
library(pheatmap)
library(viridis)

###============================================================================= parameter 
popfile <- "/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/007_ASMai/000_popInfo/003_PopInfo_APG_addArchaic.txt"
dfpop <- read_tsv(popfile) %>% select(Assembly_ID, Pop, Anc)
dfpara <- readxl::read_xlsx("/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/011_adaptiveIntrogression/005_geneHaplotype/000_gene_hsmap/gene_block_hsmap.xlsx")

base_hapmap_path <- "/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/012_haplotype_network/001_haplotype"
base_output_path <- "/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/012_haplotype_network/002_plot"
###============================================================================= parameter end

dfpara0 <- dfpara %>% 
  filter(GeneName %in% target_genes) %>% 
  mutate(a=1)

for(i in 1:nrow(dfpara0)) {
  geneName <- dfpara0$GeneName[i]
  DenBlock <- str_sub(dfpara0$ChunkID[i],2) 
  DenBlockND <- dfpara0$ChunkID[i]
  archaic <- dfpara0$Archaic[i]
  
  Start_heatmap <- dfpara0$Start_heatmap
  End_heatmap <- dfpara0$End_heatmap
  
  
  # ==================== Step 1: Data preprocessing ====================
  hap_geno <- hapSummary %>% 
    dplyr::slice(-c(1:4)) %>%  
    select(-Accession, -freq) 
  
  # Set row names as haplotype names
  rownames(hap_geno) <- hap_geno$Hap
  hap_geno <- hap_geno %>% select(-Hap)
  
  # ==================== Step 2: Calculate pairwise differences ====================
  calculate_pairwise_differences <- function(geno_df) {
    hap_names <- rownames(geno_df)
    n_haps <- length(hap_names)
    
    # Create empty matrix to store differences
    diff_matrix <- matrix(0, nrow = n_haps, ncol = n_haps,
                          dimnames = list(hap_names, hap_names))
    
    # Calculate differences between each pair of haplotypes
    for (i in 1:(n_haps-1)) {
      for (j in (i+1):n_haps) {
        hap1 <- as.character(geno_df[i, ])
        hap2 <- as.character(geno_df[j, ])
        
        # Count the number of differing sites (excluding NA)
        differences <- sum(hap1 != hap2, na.rm = TRUE)
        diff_matrix[i, j] <- differences
        diff_matrix[j, i] <- differences
      }
    }
    
    return(diff_matrix)
  }
  
  # Calculate difference matrix
  diff_matrix <- calculate_pairwise_differences(hap_geno)
  
  df_diff_matrix <- as.data.frame(diff_matrix) %>% rownames_to_column(var = "Hap")
  writexl::write_xlsx(df_diff_matrix,file.path(base_output_path,str_c(DenBlockND,"_",geneName,"_haplotype_pairwise_differences_table.xlsx")))
  # ==================== Step 3: Output difference table ====================
  # Convert to long format and save
  diff_table <- diff_matrix %>% 
    as.data.frame() %>% 
    rownames_to_column("Hap1") %>% 
    pivot_longer(cols = -Hap1, names_to = "Hap2", values_to = "Differences") %>% 
    filter(Hap1 != Hap2)  # Remove diagonal (self-comparison)
  
  # Save difference table
  write_csv(diff_table, file.path(base_output_path,str_c(DenBlockND,"_",geneName,"_haplotype_pairwise_differences.csv")))
  cat("haplotype_pairwise_differences.csv\n")
  
  # View summary statistics
  cat("Difference statistics summary:\n")
  cat("Minimum difference:", min(diff_table$Differences), "\n")
  cat("Maximum difference:", max(diff_table$Differences), "\n")
  cat("Average difference:", mean(diff_table$Differences), "\n")
  
  
  # diff_table_Archaic2Morden <- diff_table %>% 
  #   filter(Hap1 == "H070")
  
  
  # ==================== Step 4: Multiple visualization methods ====================

  pheatmap(diff_matrix,
           cluster_rows = TRUE,
           cluster_cols = TRUE,
           display_numbers = TRUE,
           number_format = "%.0f",
           number_color = "black",
           # fontsize_number = 8,
           fontsize_number = 4, ## prdm16
           color = colorRampPalette(c("blue", "white", "red"))(50),
           main = "The full matrix of pairwise differences between all the unique haplotypes",
           # filename = "haplotype_differences_clustered.pdf",
           width = 15,height = 10, ## prdm16
           # width = 10,height = 8, 
           filename = file.path(base_output_path,str_c(geneName,"_",DenBlockND,"_haplotype_differences_clustered.pdf"))
  )
  cat("haplotype_differences_clustered.pdf\n")
  
  # Network diagram visualizing difference relationships
  library(igraph)
  
  create_difference_network <- function(diff_table, threshold = 5) {
    network_edges <- diff_table %>% 
      filter(Differences <= threshold) %>% 
      select(from = Hap1, to = Hap2, weight = Differences)
    
    g <- graph_from_data_frame(network_edges, directed = FALSE)
    
    plot(g, 
         vertex.size = 15,
         vertex.color = "lightblue",
         vertex.label.cex = 0.8,
         vertex.label.color = "black",
         edge.width = 2,
         edge.color = "gray",
         main = paste("Haplotype Similarity Network (Differences ≤", threshold, ")"))
  }
  
  # pdf("haplotype_similarity_network.pdf", width = 10, height = 8)
  pdf(file.path(base_output_path,str_c(geneName,"_",DenBlock,"_differences_distribution.pdf")),width = 10, height = 8)
  create_difference_network(diff_table, threshold = 5)
  dev.off()
  cat("haplotype_similarity_network.pdf\n")
  
  # ==================== Step 5: Generate analysis report ====================
  cat("\n=== Analysis Report ===\n")
  cat("Number of haplotypes analyzed:", nrow(diff_matrix), "\n")
  cat("Number of SNP sites analyzed:", ncol(hap_geno), "\n")
  cat("Total number of pairwise comparisons:", nrow(diff_table), "\n")
  cat("Difference range:", min(diff_table$Differences), "-", max(diff_table$Differences), "sites\n")
  cat("Average genetic difference:", round(mean(diff_table$Differences), 2), "±", 
      round(sd(diff_table$Differences), 2), "sites\n")
  
  # Find haplotype pairs with minimum and maximum differences
  min_diff_pair <- diff_table[which.min(diff_table$Differences), ]
  max_diff_pair <- diff_table[which.max(diff_table$Differences), ]
  
  cat("\nMost similar haplotype pair:", min_diff_pair$Hap1, "vs", min_diff_pair$Hap2, 
      "(difference =", min_diff_pair$Differences, "sites)\n")
  cat("Most divergent haplotype pair:", max_diff_pair$Hap1, "vs", max_diff_pair$Hap2, 
      "(difference =", max_diff_pair$Differences, "sites)\n")
  
  # Save detailed report
  sink(file.path(base_output_path, str_c(geneName, "_", DenBlock, "_haplotype_analysis_report.txt")))
  cat("Haplotype Difference Analysis Report\n")
  cat("===================\n\n")
  cat("Analysis time:", format(Sys.time(), "%Y-%m-%d %H:%M:%S"), "\n\n")
  cat("Basic Statistics:\n")
  cat("- Number of haplotypes:", nrow(diff_matrix), "\n")
  cat("- Number of SNP sites:", ncol(hap_geno), "\n")
  cat("- Difference range:", min(diff_table$Differences), "-", max(diff_table$Differences), "\n")
  cat("- Average difference:", round(mean(diff_table$Differences), 2), "\n\n")
  cat("Most similar haplotype pair:", paste(min_diff_pair$Hap1, min_diff_pair$Hap2, sep = " vs "), 
      "(difference =", min_diff_pair$Differences, ")\n")
  cat("Most divergent haplotype pair:", paste(max_diff_pair$Hap1, max_diff_pair$Hap2, sep = " vs "), 
      "(difference =", max_diff_pair$Differences, ")\n")
  sink()
  
  cat("\nAnalysis complete! Generated files:\n")
  cat("1. haplotype_pairwise_differences.csv - Difference data table\n")
  cat("2. haplotype_differences_heatmap.pdf - Heatmap\n")
  cat("3. haplotype_differences_clustered.pdf - Clustered heatmap\n")
  cat("4. haplotype_differences_upper_triangle.pdf - Upper triangle heatmap\n")
  cat("5. differences_distribution.pdf - Difference distribution plot\n")
  cat("6. haplotype_similarity_network.pdf - Similarity network diagram\n")
  cat("7. haplotype_analysis_report.txt - Analysis report\n")
  
}
```

## work3: Filter haplotypes, keep the most common haplotype and Denisovan

```{r}
# Method 1: Use filter_hap function to specify haplotypes to remove
library(geneHapR)
### prdm16 >= 3 PRKCH >= 5 CSGALNACT2 >= 5 CH25H >= 5
N_fre <- 5
### Get frequency table for each haplotype
### df_hap_fre <- tibble(Hap = hapSummary$Hap, Fre = hapSummary$freq) %>% filter(str_starts(Hap,"H"))
### Get the haplotype for each sample
df_sm_hap <- tibble(Hap = hapResult$Hap, Accession = hapResult$Accession) %>% filter(str_starts(Hap,"H"))
### write_tsv(df_sm_hap,"/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/012_haplotype_network/001_test/002_sm2hap_hsmap.txt")


### Get frequency table for each haplotype
df_sm_hap2 <- df_sm_hap %>% group_by(Hap) %>% dplyr::count()

### Get haplotypes with frequency greater than or equal to n samples
df_sm_hap3 <- df_sm_hap2 %>% filter(n >= N_fre)
### Find the haplotype containing the Den sample
if(archaic == "Denisovan"){
  df_sm_hap4 <- df_sm_hap %>% filter(Accession == "Den")

}else{
  df_sm_hap4 <- df_sm_hap %>% filter(Accession == "NeanAltai")
}
### Merge filtered haplotypes and Den haplotype
df_haps_to_keep <- tibble(Hap = c(df_sm_hap4$Hap, df_sm_hap3$Hap)) 
### 
### write_tsv(df_haps_to_keep,"/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/012_haplotype_network/001_haplotype/subset/PRDM16_keepHap.txt")
### write_tsv(df_haps_to_keep,"/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/012_haplotype_network/001_haplotype/subset/PRKCH_keepHap.txt")
### write_tsv(df_haps_to_keep,"/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/012_haplotype_network/001_haplotype/subset/CSGALNACT2_keepHap.txt")
### write_tsv(df_haps_to_keep,"/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/012_haplotype_network/001_haplotype/subset/CH25H_keepHap.txt")

haps_to_remove <- df_sm_hap2 %>% anti_join(df_haps_to_keep,by=c("Hap")) %>% select(Hap) %>%  pull()

# Use filter_hap function to filter
hapResult_subset <- filter_hap(hapResult,
                          rm.mode = "haplotype",
                          haplotype.rm = haps_to_remove)

### Filter original grouping information
AccINFO_filtered <- AccINFO %>% rownames_to_column(var = "id") %>% 
  inner_join(hapResult_subset,by = c("id" = "Accession")) %>% 
  select(id,Anc) %>% 
  column_to_rownames("id")

cat("Totally ", nrow(AccINFO_filtered) -1, " id were analysis in the subset network.")
## ================================================================ Construct hapSummary
### View haplotype results
hapSummary_subset <- geneHapR::hap_summary(hapResult_subset)
# write_tsv(hapSummary,"/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/012_haplotype_network/001_test/001_hapSummary.txt")

### Visualize detailed variation information
plotHapTable(hapSummary_subset)

## ================================================================ Construct hapNet
hapNet <- get_hapNet(hapSummary_subset,
                     AccINFO = AccINFO_filtered,
                     groupName = "Anc")
```

## work4: plot geneHapR

```{r}
library(geneHapR)
###============================================================================= parameters
dfpara <- readxl::read_xlsx("/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/011_adaptiveIntrogression/005_geneHaplotype/000_gene_hsmap/gene_block_hsmap.xlsx")
# Define base path
# base_hapmap_path <- "/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/011_adaptiveIntrogression/001_DenOri"
base_output_path <- "/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/012_haplotype_network/003_plot_network"
###============================================================================= end of parameters

df_target <- dfpara %>% filter(GeneName %in% target_genes)
dfpara0 <- df_target

for(i in 1:nrow(dfpara0)) {
 # Get parameters for current gene
  geneName <- dfpara0$GeneName[i]
  DenBlock <- str_sub(dfpara0$ChunkID[i],2) 
  archaic <- dfpara0$Archaic[i]
  BlockID <- dfpara0$ChunkID[i]
  
  Start_heatmap <- dfpara0$Start_heatmap
  End_heatmap <- dfpara0$End_heatmap
  
  cat("Processing gene:", geneName, ", block:", DenBlock, "\n")
  
  ### ============================== step1: Process raw file to obtain geno dataframe

## ============================================================================= Plotting

# plot(hapNet)

par(mar = c(8, 5, 5, 5))  # mar = c(bottom, left, top, right), unit: lines
# pdf("~/Documents/geneHapR_haplotype_network_by_population.pdf", width = 12, height = 10)

plotfile <- file.path(base_output_path, str_c(geneName, "_", BlockID, "_geneHapR_haplotype_network_by_population.pdf"))
pdf(plotfile, width = 12, height = 10)

# Define population colors
population_colors <- c(AFR = "#8B4513", AMR = "#FF7F00", Archaic = "cyan", EAS = "#FFD166",
                      EUR = "#4DAF4A", SAS = "#984EA3", SEA = "#E41A1C")


# Ensure colors match the column names of hapGroup
# Get hapGroup column names (population names)
hap_group_table <- attr(hapNet, "hapGroup")
group_names <- colnames(hap_group_table)

# Arrange colors in order of population names
backGround_colors <- population_colors[group_names]


# Visualization of haploNet
plotHapNet(hapNet,
           size = "freq",        # circle size
           scale = "log10",       # scale circle with 'log10(size + 1)' or log2
           cex = 1,            # size of hap symbol
           col.link = 1,         # link colors: 1 black, 2 red, 3 green, 4 blue
           link.width = 1,       # link widths: adjust thickness
           # legend = c(1, -3), # legend position
           # legend = c(30, -20), # legend position: larger x = more right, smaller y = more down
           legend = c(10, -20), # legend position
           ############# Parameter testing
           backGround = backGround_colors,  # Key: custom group colors
           # hapGroup = hap_group_table,
           # legend = FALSE
           labels.col = "gray10",
           pie.lim = c(5,13), ### Control node size range PRKCH
           cex.legend = 0.8,
# ========== Parameters controlling mutation number display ==========
           # mutations.cex = 0.1,
           show.mutation = 3   # mutation types: one of c(0,1,2,3); 3=numbers, 2=dots, 1=lines, 0=none

)

dev.off()  
  

}
```

