---
title: "Data analysis"
author: "Aoyue Bi"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: simplex 
    highlight: haddock 
---

# Gaussian mixture models function and application, and determination of the optimal model

## step1: Plot histogram to observe density distribution first

```{r}
library(tidyverse)
library(conflicted)
conflict_prefer(dplyr::filter)

infile <- "/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/007_ASMai/013_mergefromUp/001_Den_tracts.CHM13.bed.gz"
df_asmai <- read_tsv(infile) %>% mutate(width_seg = end-start, posterior = intro_prob)

### Create grouping of samples
df_popInfoAPG <- read_tsv("/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/007_ASMai/000_popInfo/001_PopInfo_APG.txt")

### Join tables
df0 <- df_asmai %>% left_join(df_popInfoAPG, by = c("Assembly_ID"))

df <- df0 %>% 
  mutate(Denisovan = weighted_match_rate) %>% 
  ### Which segments to analyze
  filter(Denisovan > 0.30) %>% 
  filter(width_seg > 20000) %>%
  filter(posterior >= 0.9) %>% 
  filter(Assembly_ID %in% c("C088-CHA-C08-Mat","C130-CYI02-Mat","C152-CKZ06-Mat","C021-CHA-S01_2-Pat","C027-CHA-S07-Mat","HG02106_hap2")) %>%
  group_by(Assembly_ID) %>% 
  mutate(N_segment = n()) %>% 
  mutate(label = str_c("Num_seg: ",N_segment)) %>% 
  ungroup()

cat("Total number", nrow(df %>% group_by(Assembly_ID) %>% count())) 

#### Overall distribution of Denisovan ratios across all segments
p <- df %>%
  ggplot(aes(x = Denisovan)) +
  geom_histogram(aes(y = after_stat(density), fill = N_segment), 
                 # fill = NA, 
                 # fill =  "#0072B2",
                 alpha = 0.7, 
                 bins = 50) +
  # Density plot (no fill, thicker line)
  geom_density(linewidth = 0.5, adjust = 0.5, alpha = 0.99, fill = NA, color = "gray15") +
  labs(
    x = "Match to Altai Denisovan",
    y = "Density"
  ) +
  scale_fill_viridis_c() +  # Use viridis palette
  scale_x_continuous(limits = c(0,1), breaks = seq(0,1,0.1)) +
  # facet_wrap(~Assembly_ID, scales = "free", ncol = 6) +
  facet_wrap(~Assembly_ID, ncol = 3, scales = "free_y") +
  theme_classic() +
  theme(
    strip.text = element_text(face = "bold"),
    legend.position = "bottom",
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1),
    text = element_text(size = 7),
    ### Facet theme settings
    strip.background = element_blank(),  # remove title border
    strip.text.x = element_text(
      face = "bold",
      hjust = 0,  # left align
      vjust = 1,  # top align
      margin = margin(t = 0, r = 0, b = 0.2, l = 0.2)),  # adjust spacing from plot
    plot.subtitle = element_text(hjust = 0.5, color = "gray40"))
      
# Save as PNG (high resolution)
filename <- str_replace(str_c("~/Documents/","CHA-S","_Den_match_rate_distributions.pdf"), "-", "_")
# ggsave(filename = filename, plot = p, width = 15, height = 30, bg = "white", limitsize = FALSE)
# Display the plot
print(p)
ggsave(filename = filename, plot = p, width = 6, height = 4, bg = "white", limitsize = FALSE)
```

## step2: Loop to perform likelihood ratio tests – applicable to match rate detection for individual segments

This chunk is a function

```{r}
library(tidyverse)
library(mclust)
library(mixtools)
library(patchwork)
library(diptest)
library(scico)

fun_getGMM_result <- function(df, Pop_para) {

  # Assume df is your data frame and df$Denisovan is the column to analyze
  # Set population name; here "Pop" can refer to a population or individual data, it's just a placeholder
  # Pop_para <- "CZA"

  # Count the number of samples
  num_seg <- nrow(df)

  ################################## Initialize results, bic_values, fit_objects
  # Initialize results data frame
  results <- tibble(
    Pop = Pop_para,
    Num = num_seg,
    k1 = NA,
    k2 = NA,
    k3 = NA,
    p_k2k1 = NA,
    p_k3k2 = NA
  )

  # Initialize BIC values and fit objects
  bic_values <- c()
  fit_objects <- list()
  ################################## Calculate BIC, populate results, output result files

  ### For k = 1, manually compute
  mean_value <- mean(df$Denisovan)
  sd_value <- sd(df$Denisovan)
  ### Compute log-likelihood (log-likelihood of normal distribution)
  log_likelihood_k1 <- sum(dnorm(df$Denisovan, mean = mean_value, sd = sd_value, log = TRUE))
  # Compute BIC (n = sample size, k = 2 parameters: mean and sd)
  n <- num_seg
  bic_value <- -2 * log_likelihood_k1 + 2 * log(n)

  ### Append the k=1 value to the vector
  bic_values <- bic_value
  results[["k1"]] <- bic_value

  df_k1_bic <- tibble(Pop = Pop_para, Num = num_seg, BIC = bic_value, BIC_diff = NA, LRT_p_value = NA, LRT_p_value_formatted = NA)
  write_tsv(df_k1_bic, str_c("~/Documents/00/", Pop_para, "_GMM_k1BIC_LRT_sig_test.txt"))

  # Create plot
  p_k1 <- ggplot(df, aes(x = Denisovan)) +
    geom_histogram(aes(y = ..density..), bins = 50, 
                   # y = after_stat(density)
                   fill = "lightblue", alpha = 0.7) +
    geom_density(color = "blue", linewidth = 0.6, adjust = 0.5) +  # control smoothness
    scale_x_continuous(limits = c(0,1), breaks = seq(0,1,0.1)) +
    labs(x = "Match to Altai Denisovan",
         title = paste(Pop_para, ": GMM Fit with", 1, "Components"),
         # title = paste("EAS: GMM Fit with", k, "Components"),
         y = "Density") +
    theme_bw() +
    scale_color_viridis_d() +
    theme(plot.title = element_text(hjust = 0.5, size = 9),
          axis.text =  element_text(size = 9),
          axis.title = element_text(size = 9),
          legend.title = element_text(size = 9),
          legend.text = element_text(size = 9))

  p_k1
    
  ggsave(str_c("~/Documents/00/", Pop_para, "_denisovan_GMManalysis_Fit", 1, ".pdf"), p_k1, width = 4/1.3, height = 3/1.3)
  
  ################################## Method 1: mclust (automatically selects k based on BIC)
  custom_bic <- mclustBIC(df$Denisovan, G = 1:3)
  png(str_c("~/Documents/00/", Pop_para, "_GMM_BIC_mclust.png"), width = 800, height = 600, res = 120)  # set image dimensions and resolution
  plot(custom_bic, main = str_c(Pop_para, " BIC values for GMM models based mclust"))  # visualize BIC values
  dev.off()

  fit_mclust <- Mclust(df$Denisovan, G = 1:3, nstart = 50)
  optimal_k <- fit_mclust$G
  summary(fit_mclust)  # view optimal number of modes (e.g., G=2 indicates bimodal)
  cat("Optimal number of components (mclust):", optimal_k, "\n")
  dfmclust <- tibble(Pop = Pop_para, k_optimal_mclust = optimal_k)
  write_tsv(dfmclust, str_c("~/Documents/00/", Pop_para, "_GMM_BIC_mclust.txt"))

  ################################## Method 2: mixtools (remember to sort parameters)
  # 3. Loop k = 2 to k = 5 for fitting and plotting
  gmm_plots <- list()
  tukey_results <- list()

  # Loop to compute BIC values for 1 to 5 Gaussian components
  for (k_in in 2:3) { ################################## start of loop for k_in in 2:5

    set.seed(123)  # any fixed value

    # Use withCallingHandlers to catch specific warnings
    fit_k_in <- tryCatch({
      withCallingHandlers({
        fit <- normalmixEM(df$Denisovan, k = k_in, arbmean = TRUE, epsilon = 1e-3)
        
        # Check if any variance approaches zero
        if (any(fit$sigma < 1e-4)) {
          warning(paste("Population", Pop_para, "detected variance near zero at k =", k_in))
          return(NULL)
        }
        fit
      },
      warning = function(w) {
        if (grepl("One of the variances is going to zero", w$message)) {
          # Return NULL for this specific warning
          invokeRestart("muffleWarning")
          return(NULL)
        }
      })
    }, error = function(e) {
      warning(paste("Population", Pop_para, "failed to fit at k =", k_in, ":", e$message))
      NULL
    })
    
    # If fit failed or returned NULL, skip subsequent logic for this k_in
    if (is.null(fit_k_in)) {
      bic_values <- c(bic_values, NA)
      fit_objects[[k_in]] <- NULL
      results[[paste0("k", k_in)]] <- NA
      next
    }

    # Sort parameters by mean
    order_idx <- order(fit_k_in$mu)
    fit_k_in$mu <- fit_k_in$mu[order_idx]
    fit_k_in$sigma <- fit_k_in$sigma[order_idx]
    fit_k_in$lambda <- fit_k_in$lambda[order_idx]
    fit_k_in$posterior <- fit_k_in$posterior[, order_idx]
  
    # Compute BIC
    # Compute BIC (extract log-likelihood from fit object)
    bic <- function(loglik, k, n) {-2 * loglik + k * log(n)}
    bic_k_in <- bic(fit_k_in$loglik, k_in * 2 + (k_in - 1), nrow(df))
    bic_values <- c(bic_values, bic_k_in)
    fit_objects[[k_in]] <- fit_k_in
  
    # Store BIC value in results data frame
    results[[paste0("k", k_in)]] <- bic_k_in
  
    ################################## mixtools
    ### 1. Dynamically generate group labels (supports any k value)
    get_cluster_labels <- function(posterior_matrix) {
      apply(posterior_matrix, 1, which.max) %>% 
        as.factor() %>% 
        paste0("Group", .)
    }
  
    df[[paste0("cluster_k", k_in)]] <- get_cluster_labels(fit_k_in$posterior)
    
    # 5. Dynamically plot GMM fitted curve (supports any k value)
    plot_gmm_fit <- function(data, fit, k, Pop_in) { ################################## plot_gmm_fit function start, this function plots
      # Generate fitted curve data
      x <- seq(min(data), max(data), length.out = 1000)
      curve_data <- map_dfr(1:k, ~{
        tibble(
          x = x,
          y = fit$lambda[.x] * dnorm(x, fit$mu[.x], fit$sigma[.x]),
          component = paste0("μ = ", round(fit$mu[.x], 3))  # add mean label
        )
      })

      # Modify threshold calculation part, add tryCatch error handling
      thresholds <- map_dbl(1:(k-1), ~{
        tryCatch({
          uniroot(
            function(x) fit$lambda[.x] * dnorm(x, fit$mu[.x], fit$sigma[.x]) - 
              fit$lambda[.x+1] * dnorm(x, fit$mu[.x+1], fit$sigma[.x+1]),
            interval = c(fit$mu[.x], fit$mu[.x+1])
          )$root
        }, error = function(e) {
          # If error occurs, return NA and print warning
          warning(paste("Cannot compute threshold", .x, "for population", Pop_para, "with k =", k_in, ":", e$message))
          NA_real_
        })
      })

      # Filter out NA values if needed
      thresholds <- thresholds[!is.na(thresholds)]
        
      # Create plot
      p <- ggplot(df, aes(x = Denisovan)) +
        geom_histogram(aes(y = ..density..), bins = 50, 
                       # y = after_stat(density)
                       fill = "lightblue", alpha = 0.8) +
        geom_density(color = "blue", linewidth = 0.5, adjust = 0.5) +  # control smoothness
        geom_line(data = curve_data, aes(x, y, color = component), 
                  linewidth = 0.5) +
        # geom_vline(xintercept = thresholds, linetype = "dashed", 
        #            color = "gray50") +
        # Add mean markers
        geom_vline(
          xintercept = fit$mu,
          linetype = "dotted",
          color = "green",
          alpha = 0.7
        ) +
        geom_text(
          data = tibble(x = fit$mu, y = 1),
          aes(x = x, y = 0.6, label = paste("μ:", round(x, 3))),
          vjust = -0.5, 
          color = "darkgreen",
          size = 2
        ) +
        scale_color_viridis_d() +
        scale_x_continuous(limits = c(0.3,1), breaks = seq(0,1,0.1)) +
        labs(x = "Match to Altai Denisovan",
             title = paste(Pop_in, ": GMM Fit with", k, "Components"),
             y = "Density") +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5, size = 9),
              axis.text =  element_text(size = 9),
              axis.title = element_text(size = 9),
              legend.title = element_text(size = 9),
              legend.text = element_text(size = 9))
      
      p
      
      ggsave(str_c("~/Documents/00/", Pop_in, "_denisovan_GMManalysis_Fit", k, ".pdf"), p, width = 5, height = 2)
      
      return(p)
    } ################################## plot_gmm_fit function end

    # Plot GMM for current k
    # Generate and check plot for current k
    current_plot <- plot_gmm_fit(df$Denisovan, fit_k_in, k_in, Pop_para) +
      labs(title = paste(Pop_para, ": GMM Fit with ", k_in, " Components"))
    if (inherits(current_plot, "ggplot")) {
      gmm_plots[[k_in-1]] <- current_plot  # only save valid plots
    } else {
      warning(paste("Population", Pop_para, "k =", k_in, "GMM plot is invalid, not added to list"))
    }
  
    # When k > 1, perform ANOVA and Tukey tests
    # if (k_in > 1) {
    #   anova_result <- aov(as.formula(paste("Denisovan ~", paste0("cluster_k", k_in))), data = df)
    #   tukey_results[[k_in-1]] <- TukeyHSD(anova_result)[[1]] %>% 
    #     as.data.frame() %>% 
    #     rownames_to_column("Comparison") %>% 
    #     mutate(K = k_in, Pop = Pop_para) 
    # }
  
  } ################################## end of loop for k_in in 2:5

  ################################## Output df file
  # write_tsv(df, str_c("~/Documents/00/", Pop_para, "_MR_cluster.txt"))

  ################################## Combine all GMM plots (modified part)
  if (length(gmm_plots) > 0 && all(sapply(gmm_plots, inherits, "ggplot"))) {
    # Only combine and save if there are valid ggplot objects
    combined_gmm_plots <- wrap_plots(gmm_plots, ncol = 1)

    ggsave(
      str_c("~/Documents/00/", Pop_para, "_combined_gmm_fits_denisovan.pdf"), 
      plot = combined_gmm_plots, 
      width = 5, 
      height = 2 * length(gmm_plots)  # height dynamically based on number of plots
    )
  } else {
    # If empty, print warning and skip saving
    warning(paste("Population", Pop_para, "has no combinable GMM plots, skipping save"))
  }

  # 6. Combine all Tukey test results
  # if (length(tukey_results) > 0) {
  #   all_tukey_results <- bind_rows(tukey_results)
  #   write_tsv(all_tukey_results, str_c("~/Documents/00/", Pop_para, "_GMM_TukeyHSD_results_all_k.txt"))
  # }

  ################################## After loop, select best model, compute p-value (difference in BIC between previous and next k, chi-square test for p-value)
  # Select best model (model with smallest BIC), which.min returns position of minimum
  # Initialize LRT p-value vector
  lrt_p_values <- c() 

  # LRT for k=2
  # First check if fit_objects[[2]] exists
  if (length(fit_objects) >= 2 && !is.null(fit_objects[[2]])) {
    lrt_stat <- 2 * (fit_objects[[2]]$loglik - log_likelihood_k1)
    p_value <- pchisq(lrt_stat, df = 3, lower.tail = FALSE)
    results[["p_k2k1"]] <- p_value
    lrt_p_values <- c(lrt_p_values, p_value)
  } else {
    # If not exist, mark as NA
    results[["p_k2k1"]] <- NA
    lrt_p_values <- c(lrt_p_values, NA)
    warning(paste("Population", Pop_para, "fit result for k=2 does not exist, LRT cannot be computed"))
  }

  # LRT for k=3
  if (!is.null(fit_objects[[3]]) && !is.null(fit_objects[[2]])) {
    lrt_stat <- 2 * (fit_objects[[3]]$loglik - fit_objects[[2]]$loglik)
    p_value <- pchisq(lrt_stat, df = 3, lower.tail = FALSE)
    results[["p_k3k2"]] <- p_value
    lrt_p_values <- c(lrt_p_values, p_value)
  } else {
    results[["p_k3k2"]] <- NA
    lrt_p_values <- c(lrt_p_values, NA)
  }
  results$p_optimal <- results[[paste0("p_k", optimal_k_in, "k", optimal_k_in - 1)]]

  ################################### Output BIC table
  bic_df <- tibble(
    Pop = Pop_para,
    Num = num_seg,
    k = 1:3,
    BIC = bic_values,
    BIC_diff = c(NA, diff(bic_values)),  # BIC difference
    LRT_p_value = c(NA, lrt_p_values)   # LRT p-value
  ) %>% 
    mutate(
      LRT_p_value_formatted = ifelse(
        is.na(LRT_p_value), 
        NA, 
        format.pval(LRT_p_value, digits = 3)
      )
    )

  write_tsv(bic_df, str_c("~/Documents/00/", Pop_para, "_GMM_BIC_LRT_sig_test.txt"))

  ################################### Plot line graph of BIC values
  p <- ggplot(bic_df, aes(x = k, y = BIC)) +
    geom_line(color = "gray10", alpha = 0.5) +
    geom_point(color = "orange", alpha = 0.5) +
    geom_text(aes(label = paste0("ΔBIC: ", round(BIC_diff, 2), "\np: ", LRT_p_value_formatted)),
              # vjust = -1,  # adjust label position
              hjust = 0.9,
              size = 1.6,
              color = "blue") +
    labs(x = "Number of Components (k)",
         # title = str_c(Pop_para, "\nBIC Values for Gaussian Mixture Models"),
         title = str_c(Pop_para),
         y = "BIC Value by mixtools") +
    theme_bw() +
    theme(plot.title = element_text(hjust = 0.5, size = 9),
          axis.text =  element_text(size = 9),
          axis.title = element_text(size = 9))
  # Save plot to file
  # ggsave(str_c("~/Documents/00/", Pop_para, "_GMM_BIC_mixtools.png"), plot = p, width = 4, height = 3)

  if (inherits(p, "ggplot")) {
    # ggsave(str_c("~/Documents/00/", Pop_para, "_GMM_BIC_mixtools.png"), plot = p, width = 4, height = 3, dpi = 300)
    ggsave(str_c("~/Documents/00/", Pop_para, "_GMM_BIC_mixtools.pdf"), plot = p, width = 2, height = 2)
  } else {
    warning(paste("Population", Pop_para, "BIC line plot is invalid, skipping save"))
  }

  ################################## Output mu, sigma, lambda for each model
  # Initialize data frame to store GMM parameters
  gmm_params <- tibble(
    Pop = character(),
    Num_seg = integer(),
    K = integer(),
    Component = character(),
    mu = numeric(),
    sigma = numeric(),
    lambda = numeric()
  )

  ### For k=1,
  gmm_params <- gmm_params %>% add_row(
        Pop = Pop_para,
        Num_seg = num_seg,
        K = 1,
        Component = str_c("Group", 1),
        mu = round(mean_value, 3),
        sigma = round(sd_value, 3),
        lambda = 1
  )

  # # Loop over fit_objects (k=2 to k=3)
  for (k_in in 2:3) {
    fit <- fit_objects[[k_in]]
    for (comp in 1:k_in) {
      gmm_params <- gmm_params %>%
        add_row(
          Pop = Pop_para,
          Num_seg = num_seg,
          K = k_in,
          Component = str_c("Group", comp),
          mu = round(fit$mu[comp], 3),
          sigma = round(fit$sigma[comp], 3),
          lambda = round(fit$lambda[comp], 3)
        )
    }
  }

  # Output GMM parameters to file
  write_tsv(gmm_params, str_c("~/Documents/00/", Pop_para, "_GMM_meanSD.txt"))

  ################################# Above data, plot bar chart of mu, sigma, lambda for each model
  p_lambda <- ggplot(gmm_params, aes(x = factor(K), y = lambda, fill = factor(Component))) +
    geom_col(position = position_dodge(width = 0.8), width = 0.7) +  # side-by-side bars
    geom_text(
      aes(label = sprintf("μ=%.2f\nσ=%.2f\nλ=%.2f", mu, sigma, lambda)),  # display mu and sigma
      position = position_dodge(width = 0.8),
      vjust = -0.5,  # text position adjustment
      size = 2,
      color = "black"
    ) +
    scale_fill_scico_d(palette = "roma") +  # use scico roma palette
    labs(
      title = paste("GMM Component Proportions (Lambda) for Population:", unique(gmm_params$Pop)),
      x = "Number of Components (K)",
      y = "Proportion (Lambda)",
      fill = "Component"
    ) +
    scale_y_continuous(limits = c(0, 1.1 * max(gmm_params$lambda)), breaks = seq(0, 0.6, 0.1)) +
    theme_bw() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      legend.position = "right",
      panel.grid.major.x = element_blank()
    )

  # Save plot
  # ggsave(str_c("~/Documents/00/", Pop_para, "_GMM_lambda_plot.png"), plot = p_lambda, width = 8, height = 6, dpi = 300)

  if (inherits(p_lambda, "ggplot")) {  # core check: ensure it is a valid ggplot object
    ggsave(
      str_c("~/Documents/00/", Pop_para, "_GMM_lambda_plot.png"),
      plot = p_lambda,
      width = 8,
      height = 6,
      dpi = 300
    )
  } else {
    warning(paste("Population", Pop_para, "lambda plot is invalid, skipping save"))
  }

  # Explicitly return at the end of the function
  return(invisible(NULL))  # invisible ensures no return value is printed
}
```

## step3: data import and call the function

```{r}
infile <- "/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/007_ASMai/013_mergefromUp/001_Den_tracts.CHM13.bed.gz"
df_asmai <- read_tsv(infile) %>% mutate(width_seg = end - start, posterior = intro_prob)

### Create sample grouping
df_popInfoAPG <- read_tsv("/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/007_ASMai/000_popInfo/001_PopInfo_APG.txt")

### Join tables
df0 <- df_asmai %>% left_join(df_popInfoAPG, by = c("Assembly_ID"))

df <- df0 %>% 
  mutate(Denisovan = weighted_match_rate) %>% 
  ### Which segments to analyze
  filter(Denisovan > 0.30) %>% 
  ### Segment lengths: 10k, 15k, 20k, 25k
  filter(width_seg > 20000) %>%
  filter(posterior >= 0.9) %>% 
  filter(Assembly_ID %in% c("C088-CHA-C08-Mat","C130-CYI02-Mat","C152-CKZ06-Mat","C021-CHA-S01_2-Pat","C027-CHA-S07-Mat","HG02106_hap2")) %>%
  ######################################
  ### For melanesianHSAv1 overall test, also used segment length > 20 kbp
  # filter(Denisovan > 0.35) %>%
  # filter(Assembly_ID == "melanesianHSAv1") %>%
  group_by(Assembly_ID) %>% 
  mutate(N_segment = n()) %>% 
  mutate(label = str_c("Num_seg: ", N_segment)) %>% 
  ungroup()

cat("Total number", nrow(df %>% group_by(Assembly_ID) %>% count())) 

############ Call the function, finally actually used
if (TRUE) {
  df %>%
    group_by(Assembly_ID) %>%
    group_walk(~ {
      tryCatch({
        fun_getGMM_result(.x, .y$Assembly_ID)
      }, error = function(e) {
        message("Pop ", .y$Pop, " processing failed: ", e$message)
        # Write failure information to log file
        write_lines(paste(Sys.time(), "Pop", .y$Pop, "failed:", e$message),
                   "error_log.txt", append = TRUE)
      })
    })
}
```

## step4: merge results

```{r}
############################################################################# Move results and create folders

library(tidyverse)
library(fs)  # file system operations in tidyverse ecosystem
library(conflicted)
conflicted::conflicts_prefer(fs::path)
# Define base directory
base_dir <- path("~/Documents/00")

folder_names_sum <- c("summary_bic","summary_meanSD")
walk(folder_names_sum, ~ dir_create(path(base_dir, .x)))

# 1. Define the 9 folders to create
folder_names <- c(
  "plot1_bic",
  "plot2_gmm",
  "plot3_lambda",
  "plot4_bic_mclust",
  "plot5_density",
  "res1_bic",
  "res2_meanSD",
  "res3_optimal_bic_mclust",
  "res4bic_k1"
)

# 2. Define corresponding target file suffixes (in same order as folders)
file_suffixes <- c(
  "GMM_BIC_mixtools.pdf",
  "gmm_fits_denisovan.pdf",
  "GMM_lambda_plot.png",
  "GMM_BIC_mclust.png",
  "GMManalysis_Fit1.pdf",
  "GMM_BIC_LRT_sig_test.txt",
  "GMM_meanSD.txt",
  "GMM_BIC_mclust.txt",
  "GMM_k1BIC_LRT_sig_test.txt"
)

# 3. Create folders (if they don't exist)
walk(folder_names, ~ dir_create(path(base_dir, .x)))

# 4. Move files to corresponding folders
tibble(
  folder = folder_names,
  suffix = file_suffixes
) %>%
  pwalk(function(folder, suffix) {
    # Find files in 00 directory matching the suffix
    target_files <- dir_ls(base_dir, regexp = paste0(suffix, "$"))
    
    if (length(target_files) > 0) {
      # Build destination folder path
      dest_folder <- path(base_dir, folder)
      
      # Move files
      file_move(target_files, dest_folder)
      message(paste("Moved", suffix, "to", folder))
    } else {
      warning(paste("No files ending with", suffix, "found"))
    }
  })

message("All operations completed!")
############################################################################# Merge bic results

base_dir <- path("~/Documents/00")

library(purrr)
inputDir <- fs::path(base_dir, "res1_bic")
outfile <- fs::path(base_dir, "summary_bic", "001_matchRate_BIC.txt")

df_bicMerge <- dir(inputDir, full.names = TRUE) %>%
  keep(~ str_detect(basename(.x), "_GMM_BIC_LRT_sig_test.txt")) %>%
  purrr::map(~{ read_tsv(.x) }) %>%
  bind_rows()

write_tsv(df_bicMerge, outfile)

############################################################################## Merge meanSD results
inputDir <- fs::path(base_dir, "res2_meanSD")
outfile <- fs::path(base_dir, "summary_meanSD", "001_GMM_meanSD.txt")

df_bicMerge <- dir(inputDir, full.names = TRUE) %>%
  keep(~ str_detect(basename(.x), "GMM_meanSD.txt")) %>%
  purrr::map(~{ read_tsv(.x) }) %>%
  bind_rows()

write_tsv(df_bicMerge, outfile)
```

## step5: Calculate optimal BIC – directly based on p-value judgment
Function: Determine optimal k_fit based on BIC and p-values for k=1, k=2, k=3
Input: BIC for all cases for each individual
Output: BIC corresponding to optimal k_fit, and the corresponding pop anc information for each haploid

```{r}

library(tidyverse)
##################
### parameters
##################
if (FALSE) {
  inputfile <- c("/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/008_GMMfit/006_BIC4_segWidth_filter/01_10k/summary_bic/001_matchRate_BIC.txt")
  outputfile <- c("/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/008_GMMfit/006_BIC4_segWidth_filter/01_10k/summary_bic/002_matchRate_BIC_optimalK.txt")
}

if (FALSE) {
  inputfile <- c("/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/008_GMMfit/006_BIC4_segWidth_filter/02_15k/summary_bic/001_matchRate_BIC.txt")
  outputfile <- c("/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/008_GMMfit/006_BIC4_segWidth_filter/02_15k/summary_bic/002_matchRate_BIC_optimalK.txt")
}

if (FALSE) {
  inputfile <- c("/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/008_GMMfit/006_BIC4_segWidth_filter/04_20k/summary_bic/001_matchRate_BIC.txt")
  outputfile <- c("/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/008_GMMfit/006_BIC4_segWidth_filter/04_20k/summary_bic/002_matchRate_BIC_optimalK.txt")
}

if (FALSE) {
  inputfile <- c("/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/008_GMMfit/006_BIC4_segWidth_filter/03_25k/summary_bic/001_matchRate_BIC.txt")
  outputfile <- c("/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/008_GMMfit/006_BIC4_segWidth_filter/03_25k/summary_bic/002_matchRate_BIC_optimalK.txt")
}

if (TRUE) {
  inputfile <- c("/Users/Aoyue/Documents/00/summary_bic/001_matchRate_BIC.txt")
  outputfile <- c("/Users/Aoyue/Documents/00/summary_bic/002_matchRate_BIC_optimalK.txt")
}

##################
### Start calculation
##################
df_bicMerge <- read_tsv(inputfile)
# df_bicMerge <- read_tsv("/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/008_GMMfit/pluses/002_chunk_matchRate_BIC.txt")
# Process data logic
df_processed <- df_bicMerge %>%
  # 1. Filter out rows with NA BIC
  filter(!is.na(BIC)) %>%
  # 2. Group by Pop
  group_by(Pop) %>%
  # 3. Ensure each Pop group has k=1,2,3 data (if original data is incomplete, adjust accordingly)
  filter(n() == 3) %>% 
  # 4. Compute BIC differences
  # mutate(
  #   bic_diff_21 = BIC[k == 2] - BIC[k == 1],
  #   bic_diff_32 = BIC[k == 3] - BIC[k == 2]
  # ) %>%
  # 5. Determine optimal k
  group_modify(~ {
    ### Decision rule: compare p-values for k=3 vs k=2; if significant, then k_fit = 3,
    ### If not significant, compare p-values for k=2 vs k=1; if significant, k_fit = 2; if not significant, k_fit = 1

    current_data <- .x
    k_optimal <- unique(current_data$k_optimal)
    # Get relevant p-values
    p_k3 <- current_data$LRT_p_value[current_data$k == 3]
    p_k2 <- current_data$LRT_p_value[current_data$k == 2]
    
    # Decision rule
    if (p_k3 < 0.05) {
      current_data <- current_data %>%
        mutate(
          k_fit = 3,
          p_fit = p_k3
        )
    }
    else if (p_k3 >= 0.05) {
      if (!is.na(p_k2) && p_k2 < 0.05) {
        current_data <- current_data %>%
          mutate(
            k_fit = 2,
            p_fit = p_k2
          )
      } else {
        current_data <- current_data %>%
          mutate(
            k_fit = 1,
            p_fit = NA
          )
      }
    }

    # Key: must return a data frame (current_data already has new columns added)
    return(current_data)
  }) %>%
  ungroup() 

### Filter data, calculate optimal k
df_processed2 <- df_processed %>% 
  filter(k == k_fit) %>% 
  select(Assembly_ID = Pop, "Num", "BIC", "BIC_diff", "LRT_p_value", "LRT_p_value_formatted", "k_fit", "p_fit")

### Count number of fits for each k_fit
table(df_processed2$k_fit)

#################################################### Join population info
### Re-group
df_newClass <- tibble::tribble(
  ~Pop, ~Pop_Manual,
  "VIKI","KHV",
  "MYBU","MMR",
  "LALS", "LAO",
  "LALL", "LAO",
  "LALT","LAO",
  "CMTO", "KHM",
  "CMJA", "KHM",
  "CMKH", "KHM",
  "CMLA", "KHM",
  "CMBR", "KHM",
  "CMLU", "KHM",
  "CMKE", "KHM",
  "CMKA", "KHM",
  "CMKR", "KHM",
  "CMKU", "KHM",
  "CMME", "KHM",
  "CMPH", "KHM",
  "CMST", "KHM",
  "CMKL", "KHM"
)

### Classify by population
df_popInfoAPG <- read_tsv("/Users/Aoyue/project_pos/APG_archaicDNA/002_dataAnalysis/007_ASMai/000_popInfo/001_PopInfo_APG.txt") %>% 
  mutate(Anc = if_else(Anc == "ACB.ASW", "AFR", Anc)) %>% 
  mutate(Anc = if_else(Anc == "EAS-APG", "EAS", Anc)) %>% 
  mutate(Anc = if_else(Pop == "MKK", "AFR", Anc)) %>% 
  left_join(df_newClass, by = c("Pop")) %>% 
  mutate(Pop_Manual = if_else(is.na(Pop_Manual), Pop, Pop_Manual)) %>% 
  mutate(Pop1 = Pop) %>% 
  mutate(Pop = Pop_Manual) %>% 
  dplyr::select(Assembly_ID, Pop, Pop1, Anc, Num_Assembly_ID)

df_processed3 <- df_processed2 %>% 
  relocate(Assembly_ID) %>% 
  left_join(df_popInfoAPG, by = c("Assembly_ID")) 
  
############################################################################# Calculate pluses proportion
write_tsv(df_processed3, outputfile)
```



